A Study of the Effectiveness of Usage Examples in REST API Documentation

### 2017-04-20 controlled_study.tex
adds the numbers for coding agreements

\subsubsection{Data Analysis}
We analyzed the data artifact qualitatively to understand the obstacles that REST API client developers face when using an API documentation that lacks usage examples. We exported a table of data from the database behind web-based REST API explorer. For each row, the table contains the following columns: API task, participant identifier, timestamp, API trial number, I1-4, response headers and body. The values for I1-4 used by the participants were manually coded to group the participant responses into categories. The categories help us determine the types of information that API client developers need in the API documentation to perform API tasks successfully. The analysis started with an empty set of codes and new codes were introduced to describe scenarios that did not fall under  codes that were already applied. The first author applied the codes on the raw data artifact and provided a coding scheme to a co-author. The co-author applied the codes to 82/539 API calls (10\% confidence interval with 95\% confidence level) made by the participants. The two authors validated the codes against each other to provide consistency.

@@

\section{Threats to Validity}
\label{sec:threat}
  The selection of the API, the tasks, and the participants may introduce selection bias
 that there are obstacles we don't know about that could be more severe, but were not observed because the tasks did not involve anything related to these obstacles.
  The analysis of the raw data may introduce a bias. To reduce this, we have involved multiple researchers in the analysis process.

@@

\section{Conclusion}
In this paper, we presented a set of problems that experienced REST API client developers face while performing API tasks using API documentation that lacks usage examples. We identified that, without examples, REST API client developers have trouble using the correct data types, correct data formats, and required HTTP headers and request body. We've also presented empirical evidence that by adding usage examples makes it possible to reduce mistakes, improve success rate and developer satisfaction of using the API. Practitioners can follow our recommendation as a set of guidelines while documenting REST APIs to improve API usability.

>>>>

\subsubsection{Data Analysis}
We analyzed the data artifact qualitatively to understand the obstacles that REST API client developers face when using an API documentation that lacks usage examples. We exported a table of data from the web-based REST API explorer. For each row, the table contains the following columns: API task, participant identifier, timestamp, API trial number, I1-4, response headers and body. The values for I1-4 used by the participants were manually coded to group the participant responses into categories. The categories help us determine the information type that API client developers need in the API documentation to perform API tasks. The analysis started with an empty set of codes and new codes were introduced to describe scenarios that did not fall under existing codes. The first author applied the codes on the raw data artifact and provided a coding scheme comprising 11 codes to a co-author. The co-author applied the codes to 82/539 API calls (10\% confidence interval with 95\% confidence level). The two authors validated the codes against each other to provide consistency. The initial coding from the co-author mismatched 4 / 11 codes for 27 / 82 API call samples, which were resolved when the coding was repeated after a discussion.

@@

\section{Threats to Validity}
\label{sec:threat}
The selection of the API, the tasks, and the participants may introduce selection bias. It is possible that there are obstacles we don't know about that could be more severe, but were not observed because the tasks did not involve anything related to these obstacles. The analysis of the raw data may introduce a bias. To reduce this, we have involved multiple researchers in the analysis process with a repeatable coding scheme.

@@

\section{Conclusion}
In this paper, we presented a set of problems that experienced REST API client developers face while performing API tasks using API documentation that lacks usage examples. We identified that, without examples, REST API client developers have trouble using the correct data types, correct data formats, and required HTTP headers and request body. We've also presented empirical evidence that by adding usage examples makes it possible to reduce mistakes, improve success rate and developer satisfaction of using the API. REST API developers can follow our recommendation as a set of guidelines while documenting REST APIs to improve API usability. REST API documentation tool developers can leverage our findings to improve reusable tool support for software developers.
